Splints of root systems for special Lie subalgebras

### 2014-12-19 AffineLieAlgebras/Splints/special-paper.tex
More on special splints

The weight $\mu$ is minuscule if $\left<\mu,\alpha^{\vee}\right>\leq 1$ for all $\alpha\in
\Delta^{+}$.

@@

We see that case (1) contains strongly multiplicity free modules of series $A_{r}, B_{r}, C_{r}$ and
exceptional Lie algebra $G_{2}$, strongly multiplicity free modules of $A_{1}$ are included in class
(4).

What we really need is the classification of all the representations where multiplicities of all
non-zero weights are equal to one. Fortunately, such a classification was obtained in
\cite{plotkin1998visual}. It consists of multiplicity-free representations and adjoint
representations of algebras $B_{r}, C_{r}$ and $F_{4}$.

>>>>

The weight $\mu$ is minuscule if $\left<\mu,\alpha^{\vee}\right>\leq 1$ for all $\alpha\in
\Delta^{+}$. Equivalently, all the weights of the irreducible representation $L^{\mu}$ lie on Weyl
group orbit of $\mu$.  Weight is quasi-minuscule if  $\left<\mu,\alpha^{\vee}\right>\leq 2$ for all $\alpha\in
\Delta^{+}$, this means that all non-zero weights lie on single Weyl group orbit.

@@

We see that case (1) contains strongly multiplicity free modules of series $A_{r}$ and exceptional
Lie algebra $G_{2}$, class (2) contains strongly multiplicity free modules of $ B_{r}, C_{r}$ ,
strongly multiplicity free modules of $A_{1}$ are included in class (4).

(From wiki:http://en.wikipedia.org/wiki/Minuscule_representation)
Each simple Lie algebra has a unique quasi-minuscule representation that is not minuscule, and the
multiplicity of the zero weight is the number of short nodes of the Dynkin diagram.

The minuscule representations are indexed by the weight lattice modulo the root lattice, or
equivalently by irreducible representations of the center of the simply connected compact group. For
the simple Lie algebras, the dimensions of the minuscule representations are given as follows.

    An (n+1
    k) for 0 ≤ k ≤ n (exterior powers of vector representation). Quasi-minuscule: n2+2n (adjoint)
    Bn 1 (trivial), 2n (spin). Quasi-minuscule: 2n+1 (vector)
    Cn 1 (trivial), 2n (vector). Quasi-minuscule: 2n2–n–1 if n>1
    Dn 1 (trivial), 2n (vector), 2n−1 (half spin), 2n−1 (half spin). Quasi-minuscule: 2n2–n (adjoint)
    E6 1 , 27, 27. Quasi-minuscule: 78 (adjoint)
    E7 1, 56. Quasi-minuscule: 133 (adjoint)
    E8 1. Quasi-minuscule: 248 (adjoint)
    F4 1. Quasi-minuscule: 26
    G2 1. Quasi-minuscule: 7

We need to consider representations of class (4) only for $m=1,2$ since root system $\Delta_{\sfr}$
can have roots of at most two different lengths. In the case $m=1$ we get strongly multiplicity free
modules of $A_{r}$ which were discussed above. And for algebra $A_{2}$ and $m=2$ the weight diagram
of representation $L^{2\omega_{1}}_{A_{2}}$ coincides with root system of exceptional Lie algebra
$G_{2}$. This case appears in the special embedding $A_{2}\to B_{3}$. But for $A_{r}, r>2$ there are
no root systems that coincide with weight diagram of $L^{2\omega_{1}}_{A_{r}}$.

What we really need is the classification of all the representations where multiplicities of all
non-zero weights are equal to one. Fortunately, such a classification was obtained in
\cite{plotkin1998visual}. It consists of multiplicity-free representations and adjoint
representations of algebras $B_{r}, C_{r}$ and $F_{4}$.  (? check with
\cite{berenshtein1990multiplicity}).

### 2014-12-19 AffineLieAlgebras/Splints/special-presentation.tex
More on special splints

\usetheme{AnnArbor}
%\usetheme{Warsaw}

\title[]{Splints of root systems for special Lie subalgebras}

\author[]{Polina Kakin, Vladimir Lyakhovsky, \underline{Anton Nazarov}}

\institute[]{
  Department of High Energy and Elementary Particle Physics\\
  Physical faculty\\
  SPSU\\
}

@@

 $M^{(\tilde\mu)}_{\sfr(\tilde\nu)}$ - the multiplicity of the weight $\tilde\nu$ in the module $L^{\tilde\mu}_{\sfr}$ of algebra $\sfr$,\\
 $\tilde\mu=\sum m_k \tilde\omega_k$,  $\tilde\omega_k$ - fundamental weights for $\sfr$, $m_k$ - Dynkin labels of $\mu$.
\end{frame}
\section{Special subalgebras}
\begin{frame}
  \frametitle{Splint in the case of special embedding}
  \begin{Def}
    Embedding $\af \rightarrow \gf $ of semisimple Lie algebras is called "special"{ }if a root system $ \Delta _ {\af} $ of algebra $ \af $ is not a subsystem of a root system $ \Delta _ {\gf} $ of algebra $ \gf $.
  \end{Def}
  Properties of special subalgebras \cite {d}
    \begin{itemize}
    \item Orthogonal projection of algebra $g$ onto root space of subalgebra $a$ contains a root system of subalgebra $a$ 
    \item Projections of the roots of algebra $g$ are linear combinations of the roots of subalgebra $a$ with integer coefficients.
    \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Orthogonal projection}
  \begin{itemize}
  \item $\Delta'$ is orthogonal projection of the root system $\Delta_g$ onto the root space of subalgebra $a$
    \begin{itemize}
    \item $\Delta'\supset\Delta_a$
    \item $\Delta'$ contains some roots of $\Delta_a$ more than once
    \end{itemize}

  \item $\Phi_g^{(\mu)'}$ is orthogonal projection of the singular element of algebra $g$ onto the root space of subalgebra $a$ 
  \item $V_{a\rightarrow g}$ is injection fan defined as follows:

    \begin {equation}
      V_{a\rightarrow g}=\prod_{\Delta'^{+}\setminus \Delta_a^{+}}
      (1- e^{-\lambda}).
    \end {equation}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Special splint}
  \begin{itemize}
  \item $\Delta'\approx\left(\Delta_a,\Delta_s\right)$ is a special splint
  \item $\Delta_s=\Delta' \backslash \Delta_a$ is a system of vectors
    \begin{itemize}
    \item if $\Delta_s$ is not a root system then singular element $\Phi_s$ is obtained by applying injection fan $V_{a\rightarrow g}$ to the branching coefficients $g\downarrow a$.
    \end{itemize}
  \item $W_s$ is a group of reflections in hyperplanes orthogonal to the vectors in $\Delta_s$
  \item $W_{inv}$ is a group of reflections in hyperplanes orthogonal to the roots from $\Delta_g$ that already lie in the root space of subalgebra $a$.
  \end{itemize}
\end{frame}

\section{Types of special splints}
\begin{frame}
  \frametitle{Special splints: $\Delta_s\subset\Delta_a$}
  \begin{Utv}
    If $\Delta'\approx\left(\Delta_a,\Delta_s\right)$ is a special splint and $\Delta_s\subset\Delta_a$ then:

    $$
    \Phi_g^{(\mu)'}=\sum_{w\in W_{inv}} \epsilon (w) w(1+\sigma_1+\dots)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\tilde{\rho_s}}\Phi_s^{(\tilde{\mu})}),
    $$
    \begin{itemize}
      \item where $\sigma_1+\dots$ are elements of the group $W_s$ that move $\Phi_s^{(\tilde{\mu})}$ to the next Weyl chamber of subalgebra $a$. The number of elements is equal to the number of the fundamental Weyl chamber of subalgebra $a$ one needs to fill the fundamental Weyl chamber of the group $W_{inv}$.

      \item The vector $\hat\mu+\hat\rho_g$ is vector of the highest lenght in the projection of the singular element $\Phi_g^{(\mu)'}$ with positive multiplicity.
      \end{itemize}
    \label{u3}
  \end{Utv}
\end{frame}

\begin{frame}
  \frametitle{Special splints: $\Delta_a\subset\Delta_s$}
  \begin{Utv}
    If $\Delta'\approx\left(\Delta_a,\Delta_s\right)$ is a special splint and $\Delta_a\subset\Delta_s$ while $W_{inv}\subset\Delta_a$ then:

    $$
    \Phi_g^{(\mu)'}=\sum_{w\in W_{inv}} \epsilon (w) w(1+\sigma_1+\dots)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\rho_s}\Phi_s^{(\tilde{\mu})});
    $$
    if there is no $W_{inv}$ or $W_{inv}\not\subset W_a$ then:
    $$
    \Phi_g^{(\mu)'}=(2-\sum_{w\in W_a} \epsilon (w) w)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\rho_s}\Phi_s^{(\tilde{\mu})}).
    $$
    \label{u2}
  \end{Utv}
\end{frame}

\begin{frame}
  \frametitle{Special splints and branching coefficients}
  \begin{Sl}
    If $\Phi_s^{(\tilde{\mu})}$ can be put in the fundamental Weyl chamber $\bar C_a$ of subalgebra $a$ then by applying injection fan to the singular element $\Phi_s^{(\tilde{\mu})}$ one obtains $M_{s (\tilde{\nu})}^{(\tilde{\mu})}=b_{(\mu-\tilde{\mu}+\tilde{\nu})}^{(\mu)}$. If $\Phi_s^{(\tilde{\mu})}$ can not be put in $\bar C_a$ then $b_{(\mu-\tilde{\mu}+\tilde{\nu})}^{(\mu)}=N_{(\tilde{\nu})}$ where $N_{(\tilde{\nu})}$ are as follows: $$V_{a\rightarrow g}^{-1}(1+\sigma_1+\dots)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\rho_s}\Phi_s^{(\tilde{\mu})}))=\sum_{\tilde{\nu}} N_{(\tilde{\nu})}e^{\tilde{\mu}}.$$
    In both cases $\tilde{\nu}\in\bar C_a$ and the fundamental Weyl chamber $\bar C_a$ begins in the point $\tilde{\mu}-\mu$.
    \label{u31}
  \end{Sl}

@@

\begin{frame}
  \frametitle{Types of special splints}
  \begin{table}[h]
    \begin{tabular}[t]{|p{1,5em}|p{5em}|p{5em}|p{5em}|p{2em}|p{6em}|}
      \hline
      Type & $s$ & $W_a\subset W_s$ & $W_s\subset W_a$ & $W_{inv} $ & Examples \\
      \hline
      1 & root system & no & yes & yes  & $B_2\rightarrow D_3$, $G_2\rightarrow B_3$\\
      \hline
      2 & root system & yes & yes & no & $A_1\rightarrow A_1+A_1$ \\
      \hline
      3 & root system & yes & yes & no & $A_1+A_1\rightarrow D_3$ \\
      \hline
      4 & not a root system & yes & no & yes & $A_1\rightarrow A_2$, $A_1\rightarrow B_3$\\
      \hline
      5 & not a root system & yes & no & no  & $A_1\rightarrow B_2$\\
      \hline
      6 & \multicolumn{4}{|l|}{$W_a\not\subset W_s$ and $W_s\not\subset W_a$} & $B_2\rightarrow D_4$\\
      \hline
    \end{tabular}
    \label {spsp}
  \end{table}

@@

  \begin{figure}[h]
    \center{\includegraphics[width=0.4\linewidth]{drawing-2}}
    \caption{Splint $D_3\approx (A_1+A_1,B_2)$: projection of the singular element $\Phi_{D_3}^{([1,1,2])'}$ can be made from singular elements $\Phi_{B_2}^{([1,2])}$ (one of then is marked with grey colour).}
    \label{ris4}
  \end{figure}

@@

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
  \item The concept of splint was generalised on the case of special embedding.
  \item Decomposition of the projection of the singular element of splintered algebra was found for both the case when one of the stems is not a root system and the case when it is.
  \item The way to simplify calculation of branching coefficients was found for the case when both stems are root systems.
  \item All special splints were arranged into types and some of the types were thoroughly studied.
  \item The definition of singular element for the vector system that is not a root system was suggested.
  \item Together with the newly defined injection fan it might serve as the base for representation theory of such systems.
  \item The type of special splints where neither of stems contains the other requires further research.
  \end{itemize}

>>>>

\usetheme{AnnArbor}
% \usetheme{Warsaw}

\title[Special splints]{Splints of root systems for special Lie subalgebras}

\author[Kakin, Lyakhovsky, Nazarov]{Polina Kakin, Vladimir Lyakhovsky, \underline{Anton Nazarov}}

\institute[SPbSU]{
  Department of High Energy and Elementary Particle Physics\\
  Physical faculty\\
  SPSU\\
}

@@

  $M^{(\tilde\mu)}_{\sfr(\tilde\nu)}$ - the multiplicity of the weight $\tilde\nu$ in the module $L^{\tilde\mu}_{\sfr}$ of algebra $\sfr$,\\
  $\tilde\mu=\sum m_k \tilde\omega_k$,  $\tilde\omega_k$ - fundamental weights for $\sfr$, $m_k$ - Dynkin labels of $\mu$.
\end{frame}
\section{Special subalgebras}
\begin{frame}
  \frametitle{Splint in the case of special embedding}
  \begin{Def}
    Embedding $\af \rightarrow \gf $ of semisimple Lie algebras is called "special"{ }if a root system $ \Delta _ {\af} $ of algebra $ \af $ is not a subsystem of a root system $ \Delta _ {\gf} $ of algebra $ \gf $.
  \end{Def}
  To construct an embedding consider some representation $L^{\mu}_{\af}$ as a subspace of Lie algebra $\gf$. Get generators of $\af$ as linear combinations of generators of $\gf$. To reduce $\gf$-representations project $\gf$ root system to Cartan subalgebra of $\af$.

  Properties of special subalgebras \cite {d}
  \begin{itemize}
  \item Orthogonal projection of algebra $\gf$ onto root space of subalgebra $\af$ contains a root system of subalgebra $\af$ 
  \item Projections of the roots of algebra $\gf$ are linear combinations of the roots of subalgebra $\af$ with integer coefficients.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Orthogonal projection and splint}
  \begin{itemize}
  \item $\Delta'$ is orthogonal projection of the root system $\Delta_{\gf}$ onto the root space of subalgebra $\af$
    \begin{itemize}
    \item $\Delta'\supset\Delta_{\af}$
    \item $\Delta'$ contains some roots of $\Delta_{\af}$ more than once
    \end{itemize}
  \item $\Phi_{\gf}^{(\mu)'}$ is orthogonal projection of the singular element of $\gf$-representation  onto the root space of subalgebra $\af$ 
  \item $\Delta'\approx\left(\Delta_{\af},\Delta_{\sfr}\right)$ is a special splint
  \item $\Delta_{\sfr}=\Delta' \backslash \Delta_{\af}$ is a system of vectors
  \item $W_{\sfr}$ is a group of reflections in hyperplanes orthogonal to the vectors in $\Delta_{\sfr}$
  \item $W_{inv}$ is a group of reflections in hyperplanes orthogonal to the roots from $\Delta_{\gf}$ that already lie in the root space of subalgebra $\af$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Examples of root system projections}
  \begin{multicols}{4}
    \begin{figure}[h!tb]
      \includegraphics[width=1.0\linewidth]{Special-A1-A1-A3}      
      \caption{$A_1+A_1\to A_3$}
      \label{fig:a1a1a3}
    \end{figure}
    \begin{figure}[h!tb]
      \includegraphics[width=1.0\linewidth]{Special-A2-A5}      
      \caption{$A_2\to A_5$}
      \label{fig:a2a5}
    \end{figure}
    \begin{figure}[h!tb]
      \includegraphics[width=1.0\linewidth]{Special-A2-D4}
      \caption{$A_2\to D_4$}
      \label{fig:a2d4}
    \end{figure}
    \begin{figure}[h!tb]
      \includegraphics[width=1.0\linewidth]{Special-G2-B3}      
      \caption{$G_2\to B_3$}
      \label{fig:g2b3}
    \end{figure}

  \end{multicols}

@@

\section{Types of special splints}
%% \begin{frame}
%%   \frametitle{Special splints: $\Delta_s\subset\Delta_a$}
%%   \begin{Utv}
%%     If $\Delta'\approx\left(\Delta_a,\Delta_s\right)$ is a special splint and $\Delta_s\subset\Delta_a$ then:
%%     
%%     $$
%%     \Phi_g^{(\mu)'}=\sum_{w\in W_{inv}} \epsilon (w) w(1+\sigma_1+\dots)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\tilde{\rho_s}}\Phi_s^{(\tilde{\mu})}),
%%     $$
%%     \begin{itemize}
%%     \item where $\sigma_1+\dots$ are elements of the group $W_s$ that move $\Phi_s^{(\tilde{\mu})}$ to the next Weyl chamber of subalgebra $a$. The number of elements is equal to the number of the fundamental Weyl chamber of subalgebra $a$ one needs to fill the fundamental Weyl chamber of the group $W_{inv}$. 
%%       
%%     \item The vector $\hat\mu+\hat\rho_g$ is vector of the highest lenght in the projection of the singular element $\Phi_g^{(\mu)'}$ with positive multiplicity.
%%     \end{itemize}
%%     \label{u3}
%%   \end{Utv}
%% \end{frame}
%% 
%% \begin{frame}
%%   \frametitle{Special splints: $\Delta_a\subset\Delta_s$}
%%   \begin{Utv}
%%     If $\Delta'\approx\left(\Delta_a,\Delta_s\right)$ is a special splint and $\Delta_a\subset\Delta_s$ while $W_{inv}\subset\Delta_a$ then:
%%     
%%     $$
%%     \Phi_g^{(\mu)'}=\sum_{w\in W_{inv}} \epsilon (w) w(1+\sigma_1+\dots)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\rho_s}\Phi_s^{(\tilde{\mu})});
%%     $$
%%     if there is no $W_{inv}$ or $W_{inv}\not\subset W_a$ then:
%%     $$
%%     \Phi_g^{(\mu)'}=(2-\sum_{w\in W_a} \epsilon (w) w)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\rho_s}\Phi_s^{(\tilde{\mu})}).
%%     $$
%%     \label{u2}
%%   \end{Utv}
%% \end{frame}
%%

\begin{frame}
  \frametitle{Types of special splints}
  \begin{table}[h]
    \begin{tabular}[t]{|p{1,5em}|p{5em}|p{5em}|p{5em}|p{2em}|p{6em}|}
      \hline
      Type & $s$ & $W_{\af}\subset W_{\sfr}$ & $W_{\sfr}\subset W_{\af}$ & $W_{inv} $ & Examples \\
      \hline
      1 & root system & no & yes & yes  & $B_2\rightarrow D_3$, $G_2\rightarrow B_3$\\
      \hline
      2 & root system & yes & yes & no & $A_1\rightarrow A_1+A_1$ \\
      \hline
      3 & root system & yes & yes & no & $A_1+A_1\rightarrow D_3$ \\
      \hline
      4 & not a root system & yes & no & yes & $A_1\rightarrow A_2$, $A_1\rightarrow B_3$\\
      \hline
      5 & not a root system & yes & no & no  & $A_1\rightarrow B_2$\\
      \hline
      6 & \multicolumn{4}{|l|}{$W_{\af}\not\subset W_{\sfr}$ and $W_{\sfr}\not\subset W_{\af}$} & $B_2\rightarrow D_4$\\
      \hline
    \end{tabular}
    \label {spsp}
    In first three cases coincidence of branching coefficients with weight multiplicities of $\sfr$-representations takes place. 
  \end{table}
\end{frame}

%%   \begin{frame}
%%     \frametitle{Special splints and branching coefficients}
%%     \begin{Sl}
%%       If $\Phi_s^{(\tilde{\mu})}$ can be put in the fundamental Weyl chamber $\bar C_a$ of subalgebra $a$ then by applying injection fan to the singular element $\Phi_s^{(\tilde{\mu})}$ one obtains $M_{s (\tilde{\nu})}^{(\tilde{\mu})}=b_{(\mu-\tilde{\mu}+\tilde{\nu})}^{(\mu)}$. If $\Phi_s^{(\tilde{\mu})}$ can not be put in $\bar C_a$ then $b_{(\mu-\tilde{\mu}+\tilde{\nu})}^{(\mu)}=N_{(\tilde{\nu})}$ where $N_{(\tilde{\nu})}$ are as follows: $$V_{a\rightarrow g}^{-1}(1+\sigma_1+\dots)(e^{\hat\mu+\hat\rho_g-\tilde{\mu}-\rho_s}\Phi_s^{(\tilde{\mu})}))=\sum_{\tilde{\nu}} N_{(\tilde{\nu})}e^{\tilde{\mu}}.$$
%%       In both cases $\tilde{\nu}\in\bar C_a$ and the fundamental Weyl chamber $\bar C_a$ begins in the point $\tilde{\mu}-\mu$.
%%       \label{u31}
%%     \end{Sl}
%% 
%% 
%%   \end{frame}
%% 
%%

@@

  \begin{figure}[h]
    \center{\includegraphics[width=0.4\linewidth]{drawing-2}}
    \caption{Splint $D_3\approx (A_1+A_1,B_2)$: projection of the singular element $\Phi_{D_3}^{([1,1,2])'}$ can be made from singular elements $\Phi_{B_2}^{([1,2])}$ (one of them is marked with grey colour).}
    \label{ris4}
  \end{figure}

@@

\begin{frame}
  \frametitle{Towards classification of splints for special embeddings}
  \begin{itemize}
  \item Number of special embeddings is infinite: arbitrary algebra $\gf$, $\af$-representation
  \item Splints are rare
  \item We can use Dynkin diagrams with multiplicities for projections
  \item Diagrams of the same rank as $\af$ should be considered, root multiplicities less or equal to 2
  \end{itemize}
\vspace{-0.3cm}
  \begin{multicols}{4}
    \begin{figure}[h!tb]
      \includegraphics[width=1.0\linewidth]{Special-A1-A1-A3}      \\
      \includegraphics[width=0.8\linewidth]{A1A1A3}      
      \caption{$A_1+A_1\to A_3$}
      \label{fig:a1a1a3}
    \end{figure}
    \begin{figure}[h!tb]
      \includegraphics[width=1.0\linewidth]{Special-A2-A5}      \\
      \includegraphics[width=0.8\linewidth]{A2A5}      \vspace{-0.3cm}
      \caption{$A_2\to A_5$}
      \label{fig:a2a5}
    \end{figure}\par
    \begin{figure}[h!tb]
      \includegraphics[width=1.0\linewidth]{Special-A2-D4}\\
      \includegraphics[width=0.8\linewidth]{A2D4}      
      \caption{$A_2\to D_4$}
      \label{fig:a2d4}
    \end{figure}
    \begin{figure}[h!tb]
      \includegraphics[width=0.8\linewidth]{Special-G2-B3}      \\
      \includegraphics[width=0.8\linewidth]{G2B3}      \vspace{-0.2cm}
      \caption{$G_2\to B_3$}
      \label{fig:g2b3}
    \end{figure}

  \end{multicols}

\end{frame}
\section{Conclusion}

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
  \item The concept of splint was generalised on the case of special embedding.
  \item The way to simplify calculation of branching coefficients was found for the case when both stems are root systems.
  \item Some generalization of Dynkin diagrams can be useful. 
  \end{itemize}
